### 2020-01-29 sourceSummaries.tex
Update sourceSummaries.tex

\documentclass[a4paper,12pt]{article}
\begin{document}
    \title{Source Summaries}
    \author{Quintin Walters, Joshua Niemann, Connor Leavesley,\\
    Daniel Capps, Jacob Ruud}
    \date{\today}
    \maketitle
\break
\tableofcontents
\break

\break
\noindent
\section{Filizzola 2018 Security}
\subsection{Citation}

\noindent
@article\{filizzola2018security,

title=\{Security Analysis of Bluetooth Technology\},

author=\{Filizzola, Daniel and Fraser, Sean and Samsonau, Nikita\},

year=\{2018\},

organization=\{MIT\}

\noindent
\}

\subsection{Main Idea}

\noindent
Enter main idea paragraph here.

\subsection{Theory}

\noindent
Enter Theory paragraph here.

\subsection{Findings}

\noindent
Enter findings paragraph here

\subsection{Future Directions}

\noindent
Enter directions paragraph here

\end{document}