### 2020-02-06 paper/sourceSummaries/summary2.tex
folder build

\section{Summary 2}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Name of member

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\pagebreak