### 2020-02-09 paper/sourceSummaries/summary3.tex

\section{Summary 3}

\noindent
\subsection{Breaking Access Controls with BLEKey}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{baseggio2015BLEKey}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\pagebreak