### 2020-02-23 paper/sourceSummaries/summary4.tex

\section{Summary 4}

\noindent
\subsection{Title}

\subsection{{O}n {P}rivacy and {S}ecurity {C}hallenges in {S}mart {C}onnected {H}omes}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax

\bibentry{bugeja2016privacy}

\subsubsection{Main Idea}

\noindent
The authors give an overview of the privacy and security challenges directed towards the smart home domain. The authors discuss different constraints, soulution, challenges, and research issues where further investigation is required.

\subsubsection{Theory}

\noindent
Game Theory. The authors identify different attack vectors for smart homes and give some solutions to these attacks.

\subsubsection{Method}

\noindent
The authors identified different categories of security risks and then listed what a solution would have to look like in order to mitigate the security risks.

\subsubsection{Findings}

\noindent
The authors identified different security risks involved in smart homes. For the devices in smart homes they found problems with resource constraints, lack of user interface, and a need for tamper resistant devices and data. For the communication used in smart homes they found problems with heterogeneous protocols and devices going in and out of the network. The authors also found an issue with devices in smart homes not supporting dynamic patches for various reasons.

\subsubsection{Future Directions}

\noindent
The authors offer multiple future research directions. They recommend further investigation into, identity management, risk assessment methods, information flow control, and security management methods relating to smart home environments.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{An Active Man-in-the-Middle Attack on Bluetooth Devices}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{BluetoothMITMTal}

\subsubsection{Main Idea}

\noindent
The author looks to outline the main issues with BLE security. He focus on a possible architecture for a man-in-the-middle attack and a case study on a MITM'd device and an associated application.

\subsubsection{Theory}

\noindent
The author is utilizing Game Theory. He is attacking a Bluetooth connection via a MITM.

\subsubsection{Method}

\noindent
The author uses a Dax-Hub SW-28 Smart Bracelet and its associated app PowerSensor and the subject of the MITM attack. He used Kali Linux as the attacker and installed GATTacker and BtleJuice. He connected a CSR 4.0 dongle to a VM to transmit data between the app, the bracelet, and the middleman. He then uses GATTacker to intercept the BLE advertisements and copy the GATT profile of the bracelet. The author uses the GATT to simulate the device. Tal was then able to send fake to the device.

\subsubsection{Findings}

\noindent
The author found that it was far to easy to intercept device communications and take control. Due to the inherent insecurities of the BLE protocol, he was able to change data that the device displayed, as well as take control of mobile camera via the app.

\subsubsection{Future Directions}

\noindent
The author does not give any ideas as to future directions. The security issues with BLE are widely know, and Tal hopes to explain the security issues here and how to exploit them in a unique manner.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak