### 2020-03-01 paper/sourceSummaries/summary5.tex
Connor Summary 5

\section{Summary 5}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{BLE Injection-Free Attack: A Novel Attack On Bluetooth Low Energy Devices}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{jaihc19}

\subsubsection{Main Idea}

\noindent
The authors propose a new method of attacking the cryptography of BLE devices, called an injection-free attack. The authors propose this attack for when forced key renegotiation is not possible.

\subsubsection{Theory}

\noindent
The theory being used is Game Theory. The authors are attacking BLE devices in a zero sum game.

\subsubsection{Method}

\noindent
The authors use a BLE device with enough interfaces to fill the bonding list of a target device. An attacker will keep pairing new devices to the target until the bonding list fills up. Once this occurs, the next device to join will cause the target device to forget a key and negotiate a new one with the new device. Now the target device will need to renegotiate keys with legitimate devices. The attacker could carry out this attack without the packet injection needed in other attacks.

\subsubsection{Findings}

\noindent
The authors found that it is remarkably easy to attack BLE devices even without abuse the weak cryptography due to the limited resources of the devices. This method could be used by attackers to maliciously pair with devices and remove legitimate devices from the bonding list of devices.

\subsubsection{Future Directions}

\noindent
The authors believe that future research should go into automating injection free attacks and the development of defenses against such attacks.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak