### 2020-03-09 paper/wireframe/wireframe.tex
updated conclusion, preliminary suppositions

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{bibentry}
\usepackage[hyphens]{url}
\usepackage[numbib]{tocbibind}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\setlength{\marginparwidth}{2cm}

\include{references.bib}

\title{
		\vspace{-1in} 	
		\usefont{OT1}{bch}{b}{n}
		\normalfont \normalsize \textsc{CSEC-472 Authentication Paper} \\ [14pt]
		\huge A Security Analysis of the Lenel BlueDiamond \\
}

\usepackage{authblk}

\author[1]{Quintin Walters}
\author[1]{Joshua Niemann}
\author[1]{Connor Leavesley}
\author[1]{Daniel Capps}
\author[1]{Jacob Ruud}

\affil[1]{\small{Computing Security, Rochester Institute of Technology}}

\begin{document}
\maketitle
\selectlanguage{english}
\begin{abstract}

Bluetooth technology has been around since the mid 1990's. It's adoption into consumer technology has made it a mainstream protocol for those developing compatible hardware. Unfortunately, weak security in it's underlying design has also made it a target for multiple different forms of attack. This can be especially dangerous when the compatible device is used for access control. In this paper we study a specific subset of devices from Lenel's new BlueDiamond lineup, testing a variety of methods that have proven affective on similarly designed hardware. These methods included assessments of physical hardware, firmware, and connected BlueDiamond mobile app for vulnerabilities and could result in a breach of access control.

\end{abstract}

\section{Keywords}

Bluetooth

\section{Introduction}

In the last couple of years, the physical security community has seen an abundace of standard mechanical locks being replaced with \"smart locks\" designed to make accessing secure areas simpler and safer. Only to discover that these smart locks can be opened with just as little effort as picking a standard lock requires. In response to this, corporations have been rolling out revision after revision of their smart lock technology to try and combat the evolution of access control bypass.

Our research revolves around the latest in smart lock technology from Lenels BlueDiamond range. We will use a variety of previously proven methods to attempt to make an unauthorized bypass of the access control upheld be these devices. Methods including physical testing of the wiegand hardware within the device, code review of the device firmware, and analysis of the connection between the device and compatible mobile app for potential vulnerabilities.

Smart locks are the future of physical access control technology, and research from us and other likeminded individuals will ensure that the smart locks deployed in the future are free from as many known issues as possible.

\section{Background \& Significance}
The purpose of this research is to test the security properties of the Lenel Blue Diamond Readers and to uncover different vulnerabilities in them so that they can be disclosed properly and eventually patched. Lenel has over 1 million Blue Diamond readers deployed in multiple industries, the most notable one being real-estate \cite{lenelbluediamondwebsite}. This study is important because the security of these locks are very important because if someone unauthorized is able to enter through these locked doors, everything on the other side would be compromised including important data, objects, and even people. Things which if they were compromised would result in extreme loss. For our research the major issues we found fall into three categories, the firmware used on the lock, the physical aspect of the lock, and the usage of the mobile app and bluetooth communication. We plan to conduct our research by using trusted sources on the subject and testing their conclusions on Lenel Blue Diamond readers. We plan to conduct our research by using trusted sources on the subject and testing their conclusions on Lenel Blue Diamond readers. Our research has it bounds inside the Lenel Blue Diamond readers mobile app and lock itself.

\section{Related Work}

Connected to the background and significance of your study is a section of your paper devoted to a more deliberate review and synthesis of prior studies related to the research problem under investigation. The purpose here is to place your project within the larger whole of what is currently being explored, while demonstrating to your readers that your work is original and innovative. Think about what questions other researchers have asked, what methods they have used, and what is your understanding of their findings and, when stated, their recommendations.

Since a literature review is information dense, it is crucial that this section is intelligently structured to enable a reader to grasp the key arguments underpinning your proposed study in relation to that of other researchers. A good strategy is to break the literature into "conceptual categories" [themes] rather than systematically or chronologically describing groups of materials one at a time. Note that conceptual categories generally reveal themselves after you have read most of the pertinent literature on your topic so adding new categories is an on-going process of discovery as you review more studies. How do you know you've covered the key conceptual categories underlying the research literature? Generally, you can have confidence that all of the significant conceptual categories have been identified if you start to see repetition in the conclusions or recommendations that are being made.

\textbf{NOTE:} Do not shy away from challenging the conclusions made in prior research as a basis for supporting the need for your paper. Assess what you believe is missing and state how previous research has failed to adequately examine the issue that your study addresses. For more information on writing literature reviews, go to: \url{https://libguides.usc.edu/writingguide/literaturereview}.

To help frame your paper's review of prior research, consider the "five C’s" of writing a literature review:
\begin{itemize}
    \item Cite, so as to keep the primary focus on the literature pertinent to your research problem.
    \item Compare the various arguments, theories, methodologies, and findings expressed in the literature: what do the authors agree on? Who applies similar approaches to analyzing the research problem?
    \item Contrast the various arguments, themes, methodologies, approaches, and controversies expressed in the literature: describe what are the major areas of disagreement, controversy, or debate among scholars?
    \item Critique the literature: Which arguments are more persuasive, and why? Which approaches, findings, and methodologies seem most reliable, valid, or appropriate, and why? Pay attention to the verbs you use to describe what an author says/does [e.g., asserts, demonstrates, argues, etc.].
    \item Connect the literature to your own area of research and investigation: how does your own work draw upon, depart from, synthesize, or add a new perspective to what has been said in the literature?
\end{itemize}

\section{Research Design \& Methods}
To provide a well rounded analysis of the device there are three primary areas of research.  First, attacks against the physical security of the device.  Second, analysis of the firmware for vulnerabilities.  Finally, testing the application and Bluetooth implementation.

The phyisical security of the lock, particularly the magnetic antitamper switch, is what protects it from well known attacks against the Wiegand protocol.  If the physical security can be bypassed than attackers could take advantage of these other attacks to bypass the lock entirely.

Firmware analysis is necessary to discover underlying vulnerabilities.  The vulnerabilities discovered here may include hardcoded credentials, service backdoors, and poorly implemented authentication methods.  These vulnerabilites would not be obvious or easily detectible without dissecting the firmware.

The application and Bluetooth implementation are the primary means which users will be interacting with the device.  Any vulnerabilities here will be exposed to outside attackers on a daily basis.  Any attacker within 30ft of the device can snoop on the Bluetooth traffic.  The application is also downloadable on the Google Play Store for free, any vulnerabilities in it would be easily accessible to an attacker.

\subsection{Firmware Analysis}
\begin{itemize}
    \item Search for hard-coded credentials.
    \item Search for vulnerabilities such as buffer overflows, default certificates, debugging services, open ports, etc.
    \item Search for improperly implemented authentication and cryptographic methods.
    \item Search for any unpatched vulnerabilities in dependencies and base code such as BlueBorne.
\end{itemize}

Firmware analysis is a solid foundation for the research to build upon.  The firmware is the software that runs on and controls the device as a whole, vulnerabilities evident in it reduce the security of the device as a whole and can lead to vulnerabilities in other areas.  Firmware analysis is performed with automated tools at first, followed by manual analysis to confirm any discoveries.  A final pass through with manual analysis is done to spot any missed or overlooked vulnerabilities.  The methods of analysis will vary, depending on if the firmware is supplied or if it must be extracted it from the device.  If the firmware is provided then analysis can be performed directly, otherwise extraction must be performed and the compiled firmware will be analyzed.

\subsection{Physical Analysis}
\begin{itemize}
    \item Search the lock for potential physical bypass attacks.
    \item Search for issues with the magnetic anti-tamper switch.
    \item Search the lock for easy access to Wiegand protocol leads.
    \item Search for potential areas of destructive entry.
\end{itemize}

Enterprise locks often have many potential types of issues from the perspective of a physical attack.  The authors expect to find potential implementation errors in areas such as the configuration of the anti-tamper switch, the location of the connectors to the Wiegand network, or access to the locking motor from within the device.  In order to detect these issues, the authors will analyze the lock from the perspective of a typical wall-mounted installation.  From there, the authors will look into various ways to both non-destructively, and destructively trigger an unlock.  In a physical test, a vulnerability can take a lot of different forms.

\subsection{Bluetooth and Application Analysis}
\begin{itemize}
    \item Perform Generic Attribute Protocol (GATT) attacks to gather data about the device and conduct Man-in-the-Middle (MITM) attacks. 
    \item Perform known cryptographic attacks against Bluetooth Low Energy (BLE) to decrypt intercepted traffic, perform replay attacks and command injection.
    \item Search for hard-coded credentials, buffer overflows, and other application vulnerabilities. 
    \item Search for abusable edge cases in the authentication scheme.
\end{itemize}

BLE is a major part of this research. BLE is a Bluetooth protocol variant designed for low energy operations and such has a limited set of computing resources available. BLE does not need manual pairing to a device, meaning an application with the correct authentication data can pair automatically. Automated tools will gather preliminary data about the device by examining the GATT and Generic Access Profile (GAP) . The tool will copy he GATT over to an intermediary device and MITM all communications between the device and the application. With the device communication being intercepted, an Injection Free Attack can force a re-pairing and the intermediary device will intercept the Short Term Key. With the Short Term Key,  An attacker can derive the Long Term Key. They can then use the Long Term Key to decrypt communications and send commands on the behalf of the application. Automated tools will scan the application to discover hard-coded credentials and vulnerabilities. Finally, edge cases in the authentication scheme will be manually explored to see if an attacker could exploit permissions.

\section{Preliminary Suppositions \& Implications}
\begin{itemize}
    \item A lot of research appears to be out of date in the field of lock security.
    \item This research will cover Lenel/BlueDiamond's particular implementation of BLE, which to the Author's knowledge has not been tested in an academic enviornment before.
    \item This research will extend the Wiegand research of prior by attacking newer hardware with better security features.
    \item This research has the potential to potentially affect the future design of locks from the Carrier/Lenel brand, in addition to removing security vulnerabilities in future revisions of the product.
    \item This research will provide a baseline for future work in evaluating locks from different vendors.
    \item If critical vulnerabilities are discovered, these vulnerabilities could potentially be used to take over many organizations that rely on Lenel/BlueDiamond technology for access control. 
\end{itemize}

\section{Conclusions}
\begin{itemize}
    \item Go over the conclusions found in the paper.
    \item Remark on the potential implications of the findings.
    \item Express a baseline on how other researchers can start to go about testing other IOT lock technology.
    \item Document the ease of finding such vulnerabilities.
\end{itemize}

\section{Acknowledgements}
We would like to thank Lenel for access to their BlueDiamond lock and permission to conduct a test of its security.

\bibliographystyle{ieee}
\bibliography{references}

\bibentry{lenelbluediamondwebsite}

\end{document}