### 2020-03-29 paper/sourceSummaries/summary6.tex
added joshua's 6th source summary

\section{Summary 6}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{L}ock {P}icking in the {E}ra of {I}nternet of {T}hings}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{8887393}

\subsubsection{Main Idea}

\noindent
The authors perform a security analysis of a common Bluetooth Smart Lock created by a major brand.  A MasterLock was chosen because of the strong brand recognition in the lock segment, with a smaller emphasis on IOT devices.

\subsubsection{Theory}

\noindent
Game Theory.  The authors perform an attack on a common lock in order to determine inherent weaknesses in the product.

\subsubsection{Method}

\noindent
The authors conducted an analysis on the backend server API calls.  They do this by reverse engineering the mobile application for the product.  From there, valid requests were sent to determine how the API server would handle various different scenarios.

\subsubsection{Findings}

\noindent
The authors found that guest and master codes are statically generated, meaning that they could not be revoked.  The authors also found that many APIs could be misused to generate unliimited future access codes from the perspectie of a limited guest account.

\subsubsection{Future Directions}

\noindent
The authors discuss the possibility of investigating the bluetooth connection between the lock and the phone.  Although a limited amount of bluetooth research was done in the authors' testing, a lot more work can be done in investigating the bluetooth link, particularly on guest accounts.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak