### 2020-03-29 paper/sourceSummaries/summary6.tex

\section{Summary 6}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Bluetooth: With Low Energy comes Low Security}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{rmike13}

\subsubsection{Main Idea}

\noindent
The simplification of Bluetooth Low Energy to reduce power use caused a number of security issues. The author gives an overview of Bluetooth Low Energy, BLE packet injection, and encryption weaknesses.

\subsubsection{Theory}

\noindent
This paper is utilizing Game Theory. The author attacks the BLE protocol in a zero sum game.

\subsubsection{Method}

\noindent
The author implements several types of Bluetooth attacks utilizing an Ubertooth: eavesdropping, packet injection, and bypassing encryption. He then explores the flaws and devises a number of mitigations to fix the issues present.

\subsubsection{Findings}

\noindent
The author found that attacking BLE is trivial after a few, small technical hurdles. To eavesdrop on a BLE device, an attacker needs to know the hop interval, hop increment, access address, and CRC init. Ryan used this information to calculate which of the 37 channels the connection would switch to next. There are three types of methods to set up a Temporary Key (TK) for BLE, Just Works, with a key of all zeros, a six digit pin, or OOB, which is an agreed out-of-band 128 bit key. The author found that this key could be brute forced in less than a second if the device used a six digit pin. He can use the TK to derive the Short Term Key, and that to derive the Long Term Key. OOB is far harder for an attacker to crack, but exchanging a key out-of-band is difficult and not practical in practice.

\subsubsection{Future Directions}

\noindent
The author believes that future researchers should investigate main-in-the-middle attacks against Bluetooth devices to increase the effectiveness of the proposed attacks.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{L}ock {P}icking in the {E}ra of {I}nternet of {T}hings}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{8887393}

\subsubsection{Main Idea}

\noindent
The authors perform a security analysis of a common Bluetooth Smart Lock created by a major brand.  A MasterLock was chosen because of the strong brand recognition in the lock segment, with a smaller emphasis on IOT devices.

\subsubsection{Theory}

\noindent
Game Theory.  The authors perform an attack on a common lock in order to determine inherent weaknesses in the product.

\subsubsection{Method}

\noindent
The authors conducted an analysis on the backend server API calls.  They do this by reverse engineering the mobile application for the product.  From there, valid requests were sent to determine how the API server would handle various different scenarios.

\subsubsection{Findings}

\noindent
The authors found that guest and master codes are statically generated, meaning that they could not be revoked.  The authors also found that many APIs could be misused to generate unliimited future access codes from the perspectie of a limited guest account.

\subsubsection{Future Directions}

\noindent
The authors discuss the possibility of investigating the bluetooth connection between the lock and the phone.  Although a limited amount of bluetooth research was done in the authors' testing, a lot more work can be done in investigating the bluetooth link, particularly on guest accounts.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak