### 2020-04-05 paper/sourceSummaries/summary7.tex
Source 7 Dan

\section{Summary 7}

\noindent
\subsection{{A} {H}olistic {F}ramework for {O}pen {L}ow-{P}ower {I}nternet of {T}hings {T}echnology {E}cosystems}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax

\bibentry{hu2018holistic}

\subsubsection{Main Idea}

\noindent
The authors identify the working framework for low-power Iot ecosystems and find out the principle of driving the future trends in the industry and research community. The  authors also take a look at IoT technology ecosystems to find four primary business models that can lead to the proposed ecosystem framework.

\subsubsection{Theory}

\noindent
The authors used Systems Theory. They observed how IoT ecosystems work with each other to identify a framework that is holistic.

\subsubsection{Method}

\noindent
The authors found four primary business models for low-power IoT ecosystems which intersect with one another. The first is semiconductor-driven models, the second is data service driven models, the third is a vertical application-driven model, and the fourth is a framework-driven model. Using these four models the authors created a holistic framework for low-power IoT ecosystems.

\subsubsection{Findings}

\noindent
The author’s framework contains two parts in the front plane. The technical IoT building blocks and the common vertical business segments. The horizontal business segments are briefly shown on the top plane of the framework. The architecture plane indicates the various systems following an architecture based on the common IoT building blocks. On the side plane, the commercial and research maker demands are shown as the market driving force of the elements shown on the front plane. The authors concluded that AllJoyn and ARM are the successful IoT ecosystems that fit the previously proposed model and framework.

\subsubsection{Future Directions}

\noindent
Is there a way to apply this holistic framework to a real world ecosystem? How do you identify a business model in an IoT technology ecosystem to integrate these technologies?

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Extracting the Security Features Implemented in a Bluetooth LE Connection}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{zayas18}

\subsubsection{Main Idea}

\noindent
Many of the IoT devices adopting Bluetooth do not impliment the security features that Bluetooth Special Interest Group developed. To promote visability of this issue, the authors have developed an application to extract the security features of a Bluetooth device to test the security of devices.

\subsubsection{Theory}

\noindent
The authors are utilizing Systems Theory. They are observing how a device interacts with other Bluetooth devices trying to connect to it.

\subsubsection{Method}

\noindent
The authors developed an Android application for two smartphones, the Zenfone Max 3 and the OnePlus 6. They are using  Lenovo HW02 fitness tracker as the second bluetooth device. The app extracts logs from the btsnoop log created by Android devices. The app can discover security information about the connection from the logs.

\subsubsection{Findings}

\noindent
When pairing the tracker to the phones the fitness tracker sent the Long Term Key (LTK) to the tracker with no encryption, even though both are using Bluetooth 4.2. When using the app, this was not the case. Unfortunately the authors did not go into greater detail about what security features were found, only focusing on the inital transfer of the LTK.

\subsubsection{Future Directions}

\noindent
The authors believe that future research should be done in further improving the security of BLE, as it is here to stay.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{C}onfidence in {S}mart {T}oken {P}roximity: {R}elay {A}ttacks {R}evisited

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{hancke2009confidence}

\subsubsection{Main Idea}

\noindent
The authors revisit the fesibility of implementing both passive and active relay attacks against smart tokens. They also discuss the security implications should the attackers succeed. Finally, the authors discuss possible actions that device owners could take to mitigate the risk of these attacks.

\subsubsection{Theory}

\noindent
The authors use game theory to explain the red vs blue scenario, treating a win for red as a breach of information and a win for blue as witholding access control.

\subsubsection{Method}

\noindent
The authors utilize a proxy-token and proxy-reader to create a virtual clone of the authenticated users access card and relay it to the reader. This can be done by creating your own hardware or using existing tools. The authors chose to implement their own proxy reader and proxy token for this experiment.

\subsubsection{Findings}

\noindent
the authors were able to successfully perfom a relay attack against an ISO 14443A contactless system using guidelines in public literature and easily obtainable hardware. They also found that timing constraints have little to no effect against relay attacks, and although 2FA is effective if nullifies some of the advantages of smart token systems. The authors suggest using distance bounding or monitoring using a trusted interface as more effective methods against this attack, although they warn that these methods are generally more expensive or complex.

\subsubsection{Future Directions}

\noindent
The authors leave open the option for further research on methods against relay attacks since they believe that the methods they proposed could be improved upon.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak