### 2020-04-06 paper/sourceSummaries/summary7.tex

\section{Summary 7}

\noindent
\subsection{{A} {H}olistic {F}ramework for {O}pen {L}ow-{P}ower {I}nternet of {T}hings {T}echnology {E}cosystems}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax

\bibentry{hu2018holistic}

\subsubsection{Main Idea}

\noindent
The authors identify the working framework for low-power Iot ecosystems and find out the principle of driving the future trends in the industry and research community. The  authors also take a look at IoT technology ecosystems to find four primary business models that can lead to the proposed ecosystem framework.

\subsubsection{Theory}

\noindent
The authors used Systems Theory. They observed how IoT ecosystems work with each other to identify a framework that is holistic.

\subsubsection{Method}

\noindent
The authors found four primary business models for low-power IoT ecosystems which intersect with one another. The first is semiconductor-driven models, the second is data service driven models, the third is a vertical application-driven model, and the fourth is a framework-driven model. Using these four models the authors created a holistic framework for low-power IoT ecosystems.

\subsubsection{Findings}

\noindent
The author’s framework contains two parts in the front plane. The technical IoT building blocks and the common vertical business segments. The horizontal business segments are briefly shown on the top plane of the framework. The architecture plane indicates the various systems following an architecture based on the common IoT building blocks. On the side plane, the commercial and research maker demands are shown as the market driving force of the elements shown on the front plane. The authors concluded that AllJoyn and ARM are the successful IoT ecosystems that fit the previously proposed model and framework.

\subsubsection{Future Directions}

\noindent
Is there a way to apply this holistic framework to a real world ecosystem? How do you identify a business model in an IoT technology ecosystem to integrate these technologies?

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Extracting the Security Features Implemented in a Bluetooth LE Connection}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{zayas18}

\subsubsection{Main Idea}

\noindent
Many of the IoT devices adopting Bluetooth do not impliment the security features that Bluetooth Special Interest Group developed. To promote visability of this issue, the authors have developed an application to extract the security features of a Bluetooth device to test the security of devices.

\subsubsection{Theory}

\noindent
The authors are utilizing Systems Theory. They are observing how a device interacts with other Bluetooth devices trying to connect to it.

\subsubsection{Method}

\noindent
The authors developed an Android application for two smartphones, the Zenfone Max 3 and the OnePlus 6. They are using  Lenovo HW02 fitness tracker as the second bluetooth device. The app extracts logs from the btsnoop log created by Android devices. The app can discover security information about the connection from the logs.

\subsubsection{Findings}

\noindent
When pairing the tracker to the phones the fitness tracker sent the Long Term Key (LTK) to the tracker with no encryption, even though both are using Bluetooth 4.2. When using the app, this was not the case. Unfortunately the authors did not go into greater detail about what security features were found, only focusing on the inital transfer of the LTK.

\subsubsection{Future Directions}

\noindent
The authors believe that future research should be done in further improving the security of BLE, as it is here to stay.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{O}n {B}ad {R}andomness and {C}loning of {C}ontactless {P}ayment and {B}uilding {S}mart {C}ards}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{6565237}

\subsubsection{Main Idea}

\noindent
The authors describe the problems with the Mifare Classic RFID card's random number generator and examine the real life implications of bad key generation.

\subsubsection{Theory}

\noindent
Game Theory.  The authors explore the possibility of directly attacking MiFare cards, through the Random Number Generator and otherwise.

\subsubsection{Method}

\noindent
The authors review recent revelations about MiFare random number generation and describe the problem with entropy.  They then review the probability distribution of the random number generators for Mifare cards.

\subsubsection{Findings}

\noindent
The authors found that random number generators in small cards like the Mifare are often not secure.  They found that using problems in random number generators allow for very easily breaking security in many smart cards.

\subsubsection{Future Directions}

\noindent
The authors do not describe any formal future directions, but one obvious future direction would be doing research on more vendors of smart cards.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{C}onfidence in {S}mart {T}oken {P}roximity: {R}elay {A}ttacks {R}evisited

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{hancke2009confidence}

\subsubsection{Main Idea}

\noindent
The authors revisit the fesibility of implementing both passive and active relay attacks against smart tokens. They also discuss the security implications should the attackers succeed. Finally, the authors discuss possible actions that device owners could take to mitigate the risk of these attacks.

\subsubsection{Theory}

\noindent
The authors use game theory to explain the red vs blue scenario, treating a win for red as a breach of information and a win for blue as witholding access control.

\subsubsection{Method}

\noindent
The authors utilize a proxy-token and proxy-reader to create a virtual clone of the authenticated users access card and relay it to the reader. This can be done by creating your own hardware or using existing tools. The authors chose to implement their own proxy reader and proxy token for this experiment.

\subsubsection{Findings}

\noindent
the authors were able to successfully perfom a relay attack against an ISO 14443A contactless system using guidelines in public literature and easily obtainable hardware. They also found that timing constraints have little to no effect against relay attacks, and although 2FA is effective if nullifies some of the advantages of smart token systems. The authors suggest using distance bounding or monitoring using a trusted interface as more effective methods against this attack, although they warn that these methods are generally more expensive or complex.

\subsubsection{Future Directions}

\noindent
The authors leave open the option for further research on methods against relay attacks since they believe that the methods they proposed could be improved upon.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Portable RFID Bumping Device}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{dijkPortableBumping}

\subsubsection{Main Idea}

\noindent
The authors proposed that RFID cards may be vulnerable to a high speed cloning attack performed by bumping into the target.  They posit that this attach can be used to gain clones of security cards without the target's knowledge or raising suspicion.

\subsubsection{Theory}

\noindent
Game Theory.  By showing the ease of cloning the cards the authors show that they are broken and need increased security in order to stay useful.

\subsubsection{Method}

\noindent
The authors used a Proxmark 3, a LG Nexus 5 with Nethunter, a Hirose usb cable, MIFARE Classic 1K, and MIFARE Classic EV1 1K.  The authors wrote a default key to the cards and then a large number of random keys.  They then generated a large number of random keys to try and gathered a large number of nonces to perform their attack.

\subsubsection{Findings}

\noindent
The authors found that the HF Hirose Antenna increased the range to 6-8cm.  They also successfully implemented BTWA to read multiple cards.  The attacks against the cards were successefully performed against both the Classic and the Classic EV1.

\subsubsection{Future Directions}

\noindent
The authors posit that more research can be done into the maximum number of cards that can be read at one time.  They also believes that the attack framework can be extended to support more attacks, optimization can be performed on the keyspace calculation software, attempts can be made to try the attack in an unstable environment, and finally optimazation of the Proxmark firmware.

\Urlmuskip=0mu plus 1mu\relax
\pagebreak