### 2020-04-12 paper/sourceSummaries/summary8.tex
Added joshua's 8th source summary

\section{Summary 8}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{O}n {T}he {P}ower {O}f {A}ctive {R}elay {A}ttacks {U}sing {C}ustom-made {P}roxies}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{6810722}

\subsubsection{Main Idea}

\noindent
The authors use custom made hardware in addition to off-the-shelf smartphones to conduct an active relay attacks.

\subsubsection{Theory}

\noindent
Game theory.  The authors attack an existing security control by replaying wireless signals.

\subsubsection{Method}

\noindent
The authors researched different types of relay attacks.  They then investigated the strengths and weaknesses of using custom hardware compared to using off the shelf equipment like a smartphone.  Given the tight time constraints of the RFID standard, the authors also investigated the round trip time for different scenarios, such as Bluetooth and Wi-Fi.  The authors then investigated different restrictions in the standard, non-rooted Android operating system in terms of cloning cards.  Finally, they used this information to build a custom relay device.

\subsubsection{Findings}

\noindent
The authors found that it is possible to use multiple smartphones on a conventional channel such as Bluetooth or Wi-Fi to relay a RFID card over a long distance using some vulnerabilities in the RFID standard.  These vulnerabilities allow for the device to slow down the transaction rate using specially crafted packets. In addition, the authors built custom hardware that performed very well for these types of attacks.  The authors compared this custom hardware to that of the ProxMark, or OpenPICC RFID-hacking hardware projects.

\subsubsection{Future Directions}

\noindent
The authors did not specify any future directions.  However, a logical step here would be to look into different protocols that might theoretically allow for longer range.  For example, could the Author's solution allow for a relay attack over an LTE connection for longer distance?  Also looking at other, non-tested types of cards could be a possible future direction.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak