### 2020-04-12 paper/sourceSummaries/summary8.tex
added summary and fixed errors in references.bib

\section{Summary 8}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{rfid15}

\subsubsection{Main Idea}

\noindent
Many companies and people use RFID technology to secure things. The authors look to detail the basics of RFID technology and methods of attacking it.

\subsubsection{Theory}

\noindent
The authors are using Game Theory.

\subsubsection{Method}

\noindent
The authors cover the history, features, components, and standards of RFID reader technology. They then cover the basic security of these readers for authentication and encryption. Finally they cover some basic attacks against RFID readers and details some potential attack scenarios.

\subsubsection{Findings}

\noindent
The security of existing RFID solutions is not where it should be. Due to manufacturers looking to lower costs, these devices tend to lack the security features needed. Many of the RFID attacks presented are trivial to carry out and are a significant problem. The feasibility of attacks is also dependent on the use of the device. Contactless payment cards are far easier to attack than reading specific RFID identifiers.

\subsubsection{Future Directions}

\noindent
The authors provide no indication of a further direction of study.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{O}n {T}he {P}ower {O}f {A}ctive {R}elay {A}ttacks {U}sing {C}ustom-made {P}roxies}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{6810722}

\subsubsection{Main Idea}

\noindent
The authors use custom made hardware in addition to off-the-shelf smartphones to conduct an active relay attacks.

\subsubsection{Theory}

\noindent
Game theory.  The authors attack an existing security control by replaying wireless signals.

\subsubsection{Method}

\noindent
The authors researched different types of relay attacks.  They then investigated the strengths and weaknesses of using custom hardware compared to using off the shelf equipment like a smartphone.  Given the tight time constraints of the RFID standard, the authors also investigated the round trip time for different scenarios, such as Bluetooth and Wi-Fi.  The authors then investigated different restrictions in the standard, non-rooted Android operating system in terms of cloning cards.  Finally, they used this information to build a custom relay device.

\subsubsection{Findings}

\noindent
The authors found that it is possible to use multiple smartphones on a conventional channel such as Bluetooth or Wi-Fi to relay a RFID card over a long distance using some vulnerabilities in the RFID standard.  These vulnerabilities allow for the device to slow down the transaction rate using specially crafted packets. In addition, the authors built custom hardware that performed very well for these types of attacks.  The authors compared this custom hardware to that of the ProxMark, or OpenPICC RFID-hacking hardware projects.

\subsubsection{Future Directions}

\noindent
The authors did not specify any future directions.  However, a logical step here would be to look into different protocols that might theoretically allow for longer range.  For example, could the Author's solution allow for a relay attack over an LTE connection for longer distance?  Also looking at other, non-tested types of cards could be a possible future direction.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Dismantling iClass and iClass Elite}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{garciaIClass}

\subsubsection{Main Idea}

\noindent
The authors reverse engineer the protocols used in the HID iClass cards in order to find vulnerabilities in the cryptographic algorithms that they implement.  They posited that since the cards relied on Security Through Obscurity there must be vulnerabilities in the implementation.

\subsubsection{Theory}

\noindent
Game Theory, the others posit that there are vulnerabilities that inherently break the security of the cards.

\subsubsection{Method}

\noindent
The researchers purchased two readers, they discovered that the chipset used can have one block written to which will reset the write only status. They then ripped the firmware from the reader for analysis.  In there analysis of the firmware they discovered multiple vulnerabilities in the cryptographic methods used on the cards.

\subsubsection{Findings}

\noindent
The authors found that the HID iClass cards have weak keys when the three rightmosst bits of each byte are identicle, when updating the key on a card the reader sends the XOR of the old key and the new key which allows for key modification, the readers themselves have privilege escalation weaknesses within the read command, the key diversification algorithm uses DES and a proprietary hash0 which introduce a weak key diversification vulnerability, and they discovered a key recovery attack based off of the key diversification weakness.  They also found that the iClass Elite cards have a redundant key diversification weakness, a weak key-byte selection within the diversification algorithm, and a key recovery attack based off of those weaknesses.  The iClass Elite card in fact only needs 15 authentication attempts to recover the master key, something that can be done in under five seconds.

\subsubsection{Future Directions}

\noindent
The authors posit that community reviewed cryptographic algorithms would be better used in the future.  They also posit that systematic and automated model checking can be implemented in order to detect implementation weaknesses like the privilege escalation vulnerability that was discovered in the iClass cards.

\Urlmuskip=0mu plus 1mu\relax
\pagebreak