### 2020-04-19 paper/sourceSummaries/summary9.tex
summary 9 initial commit

\section{Summary 9}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Classifying RFID attacks and defenses}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{mitrokotsa09}

\subsubsection{Main Idea}

\noindent
The authors classify existing RFID attacks and discuss possible countermeasures.

\subsubsection{Theory}

\noindent
The authors are using Classification Theory. RFID attacks are being split into subgroups based on similarities and differences.

\subsubsection{Method}

\noindent
The authors classify attacks based on the layer where the attack takes place: physical layer, network-transport layer, application layer, strategic layer, and multilayer. They discuss the attacks and offer possible solutions to the issues.

\subsubsection{Findings}

\noindent
The authors found that most of the attacks possible against RFID were at the physical layer via passive and active interference methods such as sending KILL commands, replay attacks, or destruction of the tags. Multiplayer attacks were only cryptobased attacks. Network-Transport layer had the tag attacks that are more common, such as cloning and spoofing.

\subsubsection{Future Directions}

\noindent
The authors offer no indication of a future direction.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{P}ractical {B}luetooth {T}raffic {S}niffing: {S}ystems and {P}rivacy {I}mplications}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{albazrqaoe2016practical}

\subsubsection{Main Idea}

\noindent
The authors discuss privacy concerns that exist with the prevalence of bluetooth sniffing in todays iot landscape. During their discussion they propose a device of their own creation that can sniff bluetooth traffic effectively using commercially available bluetooth-compliant radios.

\subsubsection{Theory}

\noindent
Game theory is used in this paper as a method of measuring success of the new reader design; Where a win is represented by the readers ability to sniff blbluetooth traffic.

\subsubsection{Method}

\noindent
The authors improve upon the design of the Ubertooth reciever by modifying the source code to add support for adaptive hop selection and run-time clock synchronization. They also found that the firmware implementation of real-time frequency hopping needed updating to achieve the results they were aiming for.

\subsubsection{Findings}

\noindent
The authors found that their ubertooth implementation is significantly more efficiant than any previous implementation, but does not accommadate for low energy traffic. Therefore their algorithm would need to be modified in order to sniff Bluetooth Low Energy traffic as effectively as they were able to sniff normal bluetooth traffic.

\subsubsection{Future Directions}

\noindent
The authors point out that their bluetooth sniffer implementation is not configured for BLE traffic. They outline in their paper that some modifications would need to be made to their source code in order for this functionality to be added.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak