### 2020-04-25 paper/sourceSummaries/summary10.tex
summary 10 initial commit

\section{Summary 10}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Trust and Security in RFID-Based Product Authentication System}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{lehtonen07}

\subsubsection{Main Idea}

\noindent
The authors examine the trust and security of RFID authentication systems using a chain-of-trust model.

\subsubsection{Theory}

\noindent
The authors are using Classification Theory. A trust model is being constructed based on cases of misuse.

\subsubsection{Method}

\noindent
The authors first agree on a definition of authentication. They use this definition to define a several types of authentication, such as location-based authentication. The authors construct the chains of trust and they look at threats and attacks. They then look at these chains for the risk of certain paths happening.

\subsubsection{Findings}

\noindent
RFID trust chains are highly complex and there are many threats faced in each chain. These chains are an effective method of mapping out trust and the threats faced to help authentication designers better secure their RFID environment.

\subsubsection{Future Directions}

\noindent
The authors offer no indication of a future direction of research.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{Experimental} {Design}}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{kirk2012experimental}

\subsubsection{Main Idea}

\noindent
Author Roger E. Kirk explains in depth the process for designing experiments that yield the most significant and meaningful results. He builds off of work done by Sir Ronald Fisher in the early 20th century, as well as touches on some items to be wary of when undertaking this process.

\subsubsection{Theory}

\noindent
This paper aims to effectively use scientific theory to aid researchers in properly exploring their research topics.

\subsubsection{Method}

\noindent
The author analyzes numerous types of experimental design with two main goals the mind. The first being to establish a causal connection between the independent and dependent variables, and the second being to extract the maximum amount of information with the minimum expenditure of resources.

\subsubsection{Findings}

\noindent
The author presents no clear comparison data as findings to his readers, and instead leaves the decision up to his readers to interperet his data however they see fit for the benefit of their experiment.

\subsubsection{Future Directions}

\noindent
In this section of his reasearch, the author suggests no future directions to continue his work.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak