### 2020-04-26 paper/sourceSummaries/summary10.tex
Added joshua's 10th source summary

\section{Summary 10}

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Trust and Security in RFID-Based Product Authentication System}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{lehtonen07}

\subsubsection{Main Idea}

\noindent
The authors examine the trust and security of RFID authentication systems using a chain-of-trust model.

\subsubsection{Theory}

\noindent
The authors are using Classification Theory. A trust model is being constructed based on cases of misuse.

\subsubsection{Method}

\noindent
The authors first agree on a definition of authentication. They use this definition to define a several types of authentication, such as location-based authentication. The authors construct the chains of trust and they look at threats and attacks. They then look at these chains for the risk of certain paths happening.

\subsubsection{Findings}

\noindent
RFID trust chains are highly complex and there are many threats faced in each chain. These chains are an effective method of mapping out trust and the threats faced to help authentication designers better secure their RFID environment.

\subsubsection{Future Directions}

\noindent
The authors offer no indication of a future direction of research.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{R}everse {E}ngineering the {C}ommunications {P}rotocol of an {RFID} {P}ublic {T}ransportation {C}ard}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{7945583}

\subsubsection{Main Idea}

\noindent
The authors use a ProxMark 3 to investigate the feasibility of cloning a two different variants of public transport metro cards used in major cities.  Both public trasnport cards were from cities based in Spain.

\subsubsection{Theory}

\noindent
Game Theory.  The authors use weaknesses in smart card protocols to allow for easy cloning.

\subsubsection{Method}

\noindent
The authors use a ProxMark 3 RDV to actively scan for RFID cards using a long-range RFID antenna and a ProxMark connected to a laptop in a backpack.

\subsubsection{Findings}

\noindent
The authors did not find a major security issue in the RFID cards they were testing, but instead were concerned with the fact that all communication was done over cleartext channels.

\subsubsection{Future Directions}

\noindent
The authors note that Near Field Communication is partially compatible, and that an Android device may add some interest to future research.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax
\pagebreak