### 2020-04-27 paper/sourceSummaries/summary10.tex
source 10

\section{Summary 10}

\noindent
\subsection{{U}nderstanding and improving security and privacy in multi-user smart homes: a design exploration and in-home user study}

\subsubsection{Group Member}

\noindent
Daniel Capps

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax

\bibentry{zeng2019understanding}

\subsubsection{Main Idea}

\noindent
The authors explore user behaviors, needs and possible solutions to multi-user security and privacy issues in generally non-adversarial smart homes. The authors also built a prototype smart home app that includes concrete features like location based access controls, supervisory access controls, and activity notifications. The authors tested their prototype with a month long in-home user study which the authors used to re-evaluate their initial design principles and identified challenges and recommendations for smart home designers and researchers.

\subsubsection{Theory}

\noindent
The authors used Game Theory because they’re testing different smart home models against one another and using feedback of users to find the best model by balancing security with usability.

\subsubsection{Method}

\noindent
The authors beginning this paper had two research questions: “How should a smart home be designed to address multi user security and privacy challenges (in generally functional households)?” and “What security and privacy behaviors and needs do these smart home users exhibit in practice?”. The authors used initial design principles like access control flexibility, user agency, respect among users, and transparency of smart home behaviors for their smart home model. For access control flexibility the authors allowed users to set up multiple access control schemes like role-based, location-based, supervisory, and reactive access control. For transparency of smart home behaviors the authors had the users be notified of any changes to the house which include the name of the device, the change in state, and the user or process responsible for causing the change. For the user agency the authors used discovery notifications which make clear which devices were nearby and could be used.

\subsubsection{Findings}

\noindent
The authors found in the user study that found misconceptions about their principles. For access control flexibility, the authors found that while it’s important to have different access control models available to the user, having too much complexity in the system can cause users to use the minimum security level. For user agency and respect, the authors found that most of the houses opted to use social norms to support user agency and minimize conflicts instead of access control. For transparency of smart home behaviors, the authors found that they need to look further into the system as users were indifferent to the system. Overall the authors showed how complexity can be a barrier to access control use.

\subsubsection{Future Directions}

\noindent
How can we improve the usability of smart home access controls? How can we design a smart home model that takes positive household norms into account? What security mechanisms can we introduce into this smart house model for adversarial situations? How can we create a natural language based access control policy creation for smart homes?

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Trust and Security in RFID-Based Product Authentication System}

\subsubsection{Group Member}

\noindent
Connor Leavesley

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{lehtonen07}

\subsubsection{Main Idea}

\noindent
The authors examine the trust and security of RFID authentication systems using a chain-of-trust model.

\subsubsection{Theory}

\noindent
The authors are using Classification Theory. A trust model is being constructed based on cases of misuse.

\subsubsection{Method}

\noindent
The authors first agree on a definition of authentication. They use this definition to define a several types of authentication, such as location-based authentication. The authors construct the chains of trust and they look at threats and attacks. They then look at these chains for the risk of certain paths happening.

\subsubsection{Findings}

\noindent
RFID trust chains are highly complex and there are many threats faced in each chain. These chains are an effective method of mapping out trust and the threats faced to help authentication designers better secure their RFID environment.

\subsubsection{Future Directions}

\noindent
The authors offer no indication of a future direction of research.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Title}

\subsubsection{Group Member}

\noindent
Joshua Niemann

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
%\bibentry{filizzola2018security}

\subsubsection{Main Idea}

\noindent
Main Idea

\subsubsection{Theory}

\noindent
Theory

\subsubsection{Method}

\noindent
Method

\subsubsection{Findings}

\noindent
Findings

\subsubsection{Future Directions}

\noindent
Future Directions

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{{Experimental} {Design}}

\subsubsection{Group Member}

\noindent
Jacob Ruud

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{kirk2012experimental}

\subsubsection{Main Idea}

\noindent
Author Roger E. Kirk explains in depth the process for designing experiments that yield the most significant and meaningful results. He builds off of work done by Sir Ronald Fisher in the early 20th century, as well as touches on some items to be wary of when undertaking this process.

\subsubsection{Theory}

\noindent
This paper aims to effectively use scientific theory to aid researchers in properly exploring their research topics.

\subsubsection{Method}

\noindent
The author analyzes numerous types of experimental design with two main goals the mind. The first being to establish a causal connection between the independent and dependent variables, and the second being to extract the maximum amount of information with the minimum expenditure of resources.

\subsubsection{Findings}

\noindent
The author presents no clear comparison data as findings to his readers, and instead leaves the decision up to his readers to interperet his data however they see fit for the benefit of their experiment.

\subsubsection{Future Directions}

\noindent
In this section of his reasearch, the author suggests no future directions to continue his work.

\Urlmuskip=0mu plus 1mu\relax

\noindent
\subsection{Designing and implementing malicious hardeware}

\subsubsection{Group Member}

\noindent
Quintin Walters

\noindent
\subsubsection{Citation}

\Urlmuskip=0mu plus 1mu\relax
%remove % to show citation, change the name to your entry
\bibentry{kignDesigning08}

\subsubsection{Main Idea}

\noindent
The authors propose that attackers can design and attack malicious hardware to carry out attacks rather than attack the software.  They also propose that this attack vector is harder to detect and would bypass security implemented at the software level.

\subsubsection{Theory}

\noindent
Game Theory.  The authors propose and attempt an attack vector that would bypass the security of a system entirely.

\subsubsection{Method}

\noindent
The authors designed two different "Illinois Malicious Processors" to be placed on the victim hardware and they created three attacks to implement with this hardware.  The IMPs were made using modified Leon3 processors.  The first design attacked the memory of the victim hardware and their second was a complete hardware shim between the victim hardware and the victim software. They then developed a Privilege Escalation attack, a login backdoor, and a password stealer.

\subsubsection{Findings}

\noindent
The authors found that their attacks added less than 0.1 percent more logic gates to the base system and they propose that this would be even less on a larger, more complicated victim.  They also found that compared to the normal operations of the system a one time attack only increased the operation time by 1.34\% and a persistent attack only increased the operation time by 13.0\%.

\subsubsection{Future Directions}

\noindent
The authors do not propose any specific future work, however, it would be beneficial to continue to develop hardware that would not be detected via current means of supply chain attack detection.  Most notably analog power analysis, chip identification, and operation timing detection.

\Urlmuskip=0mu plus 1mu\relax
\pagebreak